Modelling writing hesitations in text writing as finite mixture process

### 2023-06-06 report/results.Rmd
add gunnexp2

$$
\begin{align}
\text{iki}_{i} & \sim \theta_\text{location[i]} \cdot \text{LogN}(\beta + \delta_\text{location[i]} + \text{participant}_i, \sigma_{e'_\text{location[i]}}^2) + \\
  & (1 - \theta_\text{location[i]}) \cdot \text{LogN}(\beta + \text{participant}_i, \sigma_{e}^2)\\
\text{where: }
		&\delta \sim \text{N}(0,1)\\
\text{constraint: } & \delta > 0\\
		& \sigma_{e'}^2 > \sigma_{e}^2
\end{align}
$$
This models takes into account two source = s of participant-specific error: (1) each participant has an individual fluent typing as in the previous models; (2) each participant has in individual hesitation frequency that differs across levels of the categorical predictor.

@@

$$
\begin{align}
\text{iki}_{i} &\sim \theta_\text{location[i]} \cdot \text{LogN}(\beta_\text{location[i]} + \delta_\text{location[i]} + \text{participant}_i, \sigma_{e'_\text{location[i]}}^2) + \\
  & (1 - \theta_\text{location[i]}) \cdot \text{LogN}(\beta_\text{location[i]} + \text{participant}_i, \sigma_{e_\text{location[i]}}^2)\\
	\text{where: } & \delta \sim \text{N}(0,1)\\
\text{constraint: } & \delta > 0\\
		& \sigma_{e'}^2 > \sigma_{e}^2
\end{align}
$$

@@

```{r keyloc, results='asis'}
tibble(`Transition type` = c("Within word", "Below word", "Before sentence"),
       Description = c("Transitions between any letter",
                       "Keypress after space followed by any letter",
                       "Keypress following a space preceding any letter"),
       Example = c("T^h^e c^a^t m^e^o^w^e^d. T^h^a^t[bsp][bsp]e^n i^t s^l^e^p^t.", 
                   "The ^cat ^meowed. That[bsp][bsp]en ^it ^slept.", 
                   "The cat meowed. ^That[bsp][bsp]en it slept.")) %>% 
  kable(caption = "Transition location classification.") %>%
  footnote("'^' marks transition location, [bsp] represents backspace. IKIs were timed to the shift keypress.") %>% 
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 12)

@@

```{r}
files <- list.files(
  str_c("../stanout/", 
  c("lift", "spl2", "spl2_shift", "plantra", "cato", "c2l1")), 
  pattern = "model.+.csv", 
  full.names = T)

table <- purrr::map_dfr(files, ~read_csv(.) %>% 
                          mutate(dataset = .x)) %>% 
  select(dataset, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(dataset, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(
     across(dataset, recode, 
                "cato" = "CATO",
                "spl2" = "SPL2", 
                "spl2_shift" = "SPL2 (shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1"),
    across(model, recode, lmm = "Unimodal log-normal",
                          lmmgaus = "Unimodal normal",
                          mogbetaconstr = "Bimodal log-normal (constrained)",
                          mogbetacontr = "Bimodal log-normal (constrained)",
                          mogbetaunconstr = "Bimodal log-normal (unconstrained)",
                          lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = dataset, values_from = contains("elpd")) %>% 
  select(Model = model, ends_with("CATO"), ends_with("C2L1"), ends_with("LIFT"), ends_with("PLanTra"), ends_with("SPL2"), ends_with("+ C)"))

@@

```{r loos, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"), 6)

kable(table, 
      caption = "Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 12))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "CATO" = 2,
                     "CL21" = 2, 
                     "LIFT" = 2,
                     "PLanTra" = 2,
                     "SPL2" = 2,
                     "SPL2 (shift + C)" = 2)) %>%
  column_spec(1, width = "8em")

@@

```{r}
files <- list.files(str_c("../stanout/", 
                          c("lift", "spl2", "plantra", "cato", "c2l1")), 
                    pattern = "mogbetaun.+.csv", 
                    full.names = T)

ps <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, recode_factor, 
                "cato" = "CATO (non-dyslexic\nunmasked)",
                "spl2" = "SPL2 (L1)", 
#                "spl2 (shift + C)" = "SPL2 (L1; shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                .ordered = TRUE)) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked")) %>% 
  pivot_wider(names_from = param, 
              values_from = value) %>% 
  unnest(cols = beta:theta) %>% 
  mutate(across(c(beta, beta2), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:theta) %>% 
  summarise(across(value, 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper),
                   .names = "{.fn}"), .by = c(data, location, name)) %>% 
  filter(name %in% c("beta", "delta", "prob"))
```

@@

```{r crossstudypost, fig.cap="Across studies. Posterior parameter distribution"}
posd <- position_dodge(.65)
dotsize <- 2.5
grouplabel <- "Data set"
beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, 
            alpha = .35, 
            position = posd) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1, 4, 7, 8, 9, 11)) +
  scale_colour_colorblind() +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = grouplabel,
       shape = grouplabel) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1, 4, 7, 8, 9, 11)) +
  scale_colour_colorblind() +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = c(1, 4, 7, 8, 9, 11)) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

@@

```{r loceffect, results='asis'}
files <- list.files(str_c("../stanout/", 
                          c("lift", "spl2", "spl2_shift", "plantra", "cato", "c2l1")), 
                    pattern = "mogbetaun.+.csv", 
                    full.names = T)

ps_loc_diffs <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, recode_factor, 
                "cato" = "CATO (non-dyslexic\nunmasked)",
                "spl2" = "SPL2 (L1)", 
                "spl2_shift" = "SPL2 (L1; shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                .ordered = TRUE)) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked"),
         param %in% c("beta", "delta", "theta")) %>%
  select(param, location, data, value) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, .))) %>% 
  pivot_wider(names_from = location, values_from = value) %>%
  unnest(-c(data, param)) %>% 
  mutate(diff.1 = `before sentence` - `before word`,
         diff.2 = `before word` - `within word`) %>% 
  summarise(across(starts_with("diff"), 
                   list(mean = mean, lower = lower, upper = upper, BF = BF)),
            .by = c(data, param))

@@

tmp %>% knitr::kable(align =c("l", "l", rep("r", 6)), 
                     col.names = c("Data set", "Difference", rep(c("Est. with 95% PIs", "BF"), 3)),
                     caption = "Effect of transition location on keystroke intervals. Differences between transition locations are shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.") %>%
  kableExtra::collapse_rows(columns = 1) %>%
  kable_styling(position = "center", full_width = T, font_size = 11) %>%
  footnote("PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.", threeparttable = T) %>% 
  add_header_above(c(" " = 2, 
                     "Fluent transitions" = 2, 
                     "Slowdown for hesitations" = 2, 
                     "Probability of hesitations" = 2))

@@

>>>>

$$
\begin{align}
\text{iki}_{i} & \sim \theta_\text{location[i]} \cdot \text{LogN}(\beta + \delta_\text{location[i]} + \text{participant}_i, \sigma_{e'_\text{location[i]}}^2) + \\
  & (1 - \theta_\text{location[i]}) \cdot \text{LogN}(\beta + \text{participant}_i, \sigma_{e_\text{location[i]}}^2)\\
\text{where: }
		&\delta \sim \text{N}(0,1)\\
 	 & \text{participant} \sim \text{N}(0, \sigma_\text{p}^2) \\
\text{constraint: } & \delta > 0\\
		& \sigma_{e'}^2 > \sigma_{e}^2
\end{align}
$$
This models takes into account two source = s of participant-specific error: (1) each participant has an individual fluent typing as in the previous models; (2) each participant has in individual hesitation frequency that differs across levels of the categorical predictor.

@@

$$
\begin{align}
\text{iki}_{i} &\sim \theta_\text{location[i]} \cdot \text{LogN}(\beta_\text{location[i]} + \delta_\text{location[i]} + \text{participant}_i, \sigma_{e'_\text{location[i]}}^2) + \\
  & (1 - \theta_\text{location[i]}) \cdot \text{LogN}(\beta_\text{location[i]} + \text{participant}_i, \sigma_{e_\text{location[i]}}^2)\\
	\text{where: } & \delta \sim \text{N}(0,1)\\
	 	 & \text{participant} \sim \text{N}(0, \sigma_\text{p}^2) \\
\text{constraint: } & \delta > 0\\
		& \sigma_{e'}^2 > \sigma_{e}^2
\end{align}
$$

@@

```{r keyloc, results='asis'}
tibble(`Transition type` = c("Within word", "Below word", "Before sentence"),
       Description = c("Transitions between any letter",
                       "Keypress after space followed by any letter",
                       "Keypress following a space preceding any letter"),
       Example = c("T\\^h\\^e c\\^a\\^t m\\^e\\^o\\^w\\^e\\^d. T\\^h\\^a\\^t[bsp][bsp]e\\^n i\\^t s\\^l\\^e\\^p\\^t.", 
                   "The \\^cat \\^meowed. That[bsp][bsp]en \\^it \\^slept.", 
                   "The cat meowed. \\^That[bsp][bsp]en it slept.")) %>% 
  kable(caption = "Transition location classification.") %>%
  footnote("'\\^' marks transition location, [bsp] represents backspace. IKIs were timed to the shift keypress.") %>% 
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 12)

@@

```{r}
files <- list.files(
  str_c("../stanout/", 
  c("lift", "spl2", "spl2_shift", "plantra", "cato", "c2l1", "gunnexp2")), 
  pattern = "model.+.csv", 
  full.names = T)

table <- purrr::map_dfr(files, ~read_csv(.) %>% 
                          mutate(dataset = .x)) %>% 
  select(dataset, model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  mutate(across(dataset, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(where(is.numeric), round, 0),
         across(where(is.numeric), format, big.mark = ","),
         across(everything(), str_trim)) %>%
  transmute(
     across(dataset, recode, 
                "cato" = "CATO",
                "spl2" = "SPL2", 
                "spl2_shift" = "SPL2 (shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                "gunnexp2" = "GUNNEXP2"),
    across(model, recode, lmm = "Unimodal log-normal",
                          lmmgaus = "Unimodal normal",
                          mogbetaconstr = "Bimodal log-normal (constrained)",
                          mogbetacontr = "Bimodal log-normal (constrained)",
                          mogbetaunconstr = "Bimodal log-normal (unconstrained)",
                          lmmuneqvar = "Unimodal log-normal (unequal variance)"),
            across(elpd_diff, str_c, " (", se_diff, ")"),
            across(elpd_loo, str_c, " (", se_elpd_loo, ")")) %>% 
  mutate(across(elpd_diff, str_replace, "0 \\(0\\)", "--")) %>% 
  pivot_wider(names_from = dataset, values_from = contains("elpd")) %>% 
  select(Model = model, ends_with("GUNNEXP2"), ends_with("CATO"), ends_with("C2L1"), ends_with("LIFT"), ends_with("PLanTra"), ends_with("SPL2"), ends_with("+ C)"))

@@

```{r loos, results='asis'}
names(table)[-1] <- rep(c("$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$"), 8)

kable(table, 
      caption = "Model comparisons. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 14))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>%
  add_header_above(c(" " = 1, 
                     "GUNNEXP2" = 2,
                     "CATO" = 2,
                     "CL21" = 2, 
                     "LIFT" = 2,
                     "PLanTra" = 2,
                     "SPL2" = 2,
                     "SPL2 (shift + C)" = 2)) %>%
  column_spec(1, width = "8em")

@@

```{r}
files <- list.files(str_c("../stanout/", 
                          c("lift", "spl2", "plantra", "cato", "c2l1", "gunnexp2")), 
                    pattern = "mogbetaun.+.csv", 
                    full.names = T)

ps <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, recode_factor, 
                "cato" = "CATO (non-dyslexic\nunmasked)",
                "gunnexp2" = "GUNNEXP2 (unmasked)",
                "spl2" = "SPL2 (L1)", 
#                "spl2 (shift + C)" = "SPL2 (L1; shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                .ordered = TRUE)) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked"),
         !(str_detect(data, "GUNNEXP2") & xn == "masked")) %>% 
  pivot_wider(names_from = param, 
              values_from = value) %>% 
  unnest(cols = beta:theta) %>% 
  mutate(across(c(beta, beta2), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  pivot_longer(beta:theta) %>% 
  summarise(across(value, 
                   list(mean = mean, 
                        lower = lower, 
                        upper = upper),
                   .names = "{.fn}"), 
            .by = c(data, location, name)) %>% 
  filter(name %in% c("beta", "delta", "prob"))
```

@@

```{r crossstudypost, fig.cap="Across studies. Posterior parameter distribution"}
posd <- position_dodge(.65)
dotsize <- 2.5
grouplabel <- "Data set"
shapes <- c(1, 4, 6, 7, 8, 9, 11)
beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, 
            alpha = .35, 
            position = posd) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Posterior estimate with 95% PIs",
       x = "",
       colour = grouplabel,
       shape = grouplabel) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = data,
             group = interaction(data))) +
  geom_line(size = .5, alpha = .35, position = posd) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = data), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title.y = element_blank()) +
  guides(colour=guide_legend(nrow=2,byrow=TRUE),
         shape = guide_legend(nrow=2,byrow=TRUE))

@@

```{r loceffect, results='asis'}
files <- list.files(str_c("../stanout/", 
                          c("lift", "spl2", "spl2_shift", "gunnexp2", "plantra", "cato", "c2l1")), 
                    pattern = "mogbetaun.+.csv", 
                    full.names = T)

ps_loc_diffs <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, recode_factor, 
                "cato" = "CATO (non-dyslexic\nunmasked)",
                "spl2" = "SPL2 (L1)", 
                "gunnexp2" = "GUNNEXP2 (unmasked)",
                "spl2_shift" = "SPL2 (L1; shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                .ordered = TRUE)) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES"),
         !(str_detect(data, "CATO") & group == "dyslexic"),
         !(str_detect(data, "CATO") & task == "masked"),
         !(str_detect(data, "GUNNEXP2") & xn == "masked"),
         param %in% c("beta", "delta", "theta")) %>%
  select(param, location, data, value) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, .))) %>% 
  pivot_wider(names_from = location, values_from = value) %>%
  unnest(-c(data, param)) %>% 
  mutate(diff.1 = `before sentence` - `before word`,
         diff.2 = `before word` - `within word`) %>% 
  summarise(across(starts_with("diff"), 
                   list(mean = mean, lower = lower, upper = upper, BF = BF)),
            .by = c(data, param))

@@

tmp %>% knitr::kable(align =c("l", "l", rep("r", 6)), 
                     col.names = c("Data set", "Difference", rep(c("Est. with 95% PIs", "BF"), 3)),
                     caption = "Effect of transition location on keystroke intervals. Differences between transition locations are shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.") %>%
  kableExtra::collapse_rows(columns = 1) %>%
  kable_styling(position = "center", full_width = T, font_size = 11) %>%
  footnote("PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.", threeparttable = T) %>% 
  add_header_above(c(" " = 2, 
                     "Fluent transitions" = 2, 
                     "Slowdown for hesitations" = 2, 
                     "Probability of hesitations" = 2)) %>% 
  column_spec(1, width = "3cm")

@@

### Unconstrained mixture model

#### Posterior parameter estimates

```{r gunnpostun, fig.cap="GUNNEXP2 (unconstrained model). Posterior parameter distribution"}
file <- list.files(path = "../stanout/gunnexp2", 
                   pattern = "mogbetaun.*.csv$",
                   full.names = T)

ps <- read_csv(file) %>% 
  filter(param %in% c("beta", "beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>%
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(location, xn, name)) %>% 
  mutate(#group = str_c("Topic: ",topic, "; ", xn))
         group = str_c(xn))

posd <- position_dodge(.35)
dotsize <- 3
#shapes <- 1:length(unique(ps$group))
#shapes <- c(2, 7, 8, 9)
shapes <- c(8, 9)

beta <- filter(ps, name == "beta") %>% 
  mutate(name = beta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = group)) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = group), 
             size = dotsize, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank())

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = group)) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = group), 
             size = dotsize, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank())

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = group)) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = group), 
             size = dotsize, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank())

#### Masking effect

```{r shiftcellmeansgunn, results = 'asis'}
# cell means and data set difference
cellmeans <- ps %>% 
  select(-group) %>% 
  mutate(across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(c(ends_with("beta"),
                  ends_with("delta")), 
                ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(xn, location), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = xn, 
              values_from = est)

tmp <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  pivot_wider(names_from = xn, values_from = value) %>% 
  unnest(c(masked, unmasked)) %>% 
  mutate(diff = masked - unmasked) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, recode, theta = "prob")) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location) %>% 
  select(param, 
         `Transition location` = location, 
         Unmasked = unmasked,
         Masked = masked,
         "Difference" = mean, BF)

write_csv(tmp, "tables/masking_effect_cellmeans.csv")

tmp[,-1] %>% knitr::kable(align = c("l", "r", "r", "r", "r"), 
                          caption = "Mixture model estimates for key transitions. Cell means are shown for the masked and unmasked writing task in msecs for fluent key-transitions, the slowdown for long transitions and the probability of disfluent transitions. The effect for masking is shown on log scale (for transition durations) and logit scale for probability of disfluent transitions. 95% PIs in brackets.") %>%
  #  cell_spec(1, 3:4) %>% 
  kableExtra::collapse_rows(columns = 1) %>%
  kable_styling(position = "center", full_width = T) %>%
  footnote("PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.", 
           threeparttable = T) %>% 
  group_rows("Fluent transitions", 1, 3) %>%
  group_rows("Disfluencies", 4, 6) %>%
  group_rows("Probability of disfluencies", 7, 9)

### Cononstrained mixture model

#### Posterior parameter estimates

```{r}
file <- list.files(path = "../stanout/gunnexp2", 
                   pattern = "mogbetac.*.csv$",
                   full.names = T)

ps_beta <- read_csv(file) %>% 
  filter(param %in% c("beta")) %>% 
  nest(beta = value) %>% 
  select(beta)

beta <- ps_beta %>% 
   unnest(beta) %>% 
   mutate(across(value, exp)) %>% 
   summarise(across(value, list(mean = mean, 
                                lower = lower, 
                                upper = upper), .names = "{.fn}")) %>% 
  mutate(across(everything(), round, 0)) %>% 
  transmute(result = str_c(mean, " msecs, PI: (", lower, ", ", upper, ")" )) %>% 
  pull(result)
```

The posterior of the constrained model is shown in Figure \@ref(fig:gunnpost) showing the posterior slowdown for disfluent keystrokes (left panel) and the probability of disfluent keystrokes (right panel). Fluent keystroke transitions are distributed around a posterior mean of `r beta`.

```{r gunnpost, fig.cap="GUNNEXP2 (constrained model). Posterior parameter distribution."}

ps <- read_csv(file) %>% 
  filter(param %in% c("beta2", "prob")) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  bind_cols(ps_beta) %>% 
  unnest(cols = c(beta, beta2, prob)) %>% 
  rename(beta = value) %>% 
  mutate(across(c(beta2, beta), exp),
         delta = beta2 - beta,
         across(location, str_replace, " ", "\n")) %>% 
  select(-beta2) %>% 
  pivot_longer(c(beta, prob, delta)) %>%
  filter(name != "beta") %>% 
  summarise(across(value, list(mean = mean, 
                               lower = lower, 
                               upper = upper), .names = "{.fn}"),
            .by = c(location, xn, name)) %>% 
  mutate(#group = str_c("Topic: ",topic, "; ", xn))
         group = str_c(xn))

posd <- position_dodge(.35)
dotsize <- 3
#shapes <- 1:length(unique(ps$group))
shapes <- c(8, 9)

delta <- filter(ps, name == "delta") %>% 
  mutate(name = delta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = group)) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = group), 
             size = dotsize, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "",
       x = "Transition location",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank()) 
#  guides(colour=guide_legend(nrow=2,byrow=F),
#         shape = guide_legend(nrow=2,byrow=F))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = location,
             shape = group)) +
  geom_errorbar(width = .2, alpha = .75, position = posd) +
  geom_point(aes(colour = group), 
             size = dotsize, position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(y = "",
       x = "Transition location",
       colour = "",
       shape = "") +
  theme(axis.title.y = element_blank())

delta + theta + plot_layout(guides = "collect")

#### Masking effect

```{r shiftcellmeansgunn2, results = 'asis'}
# cell means and data set difference
cellmeans <- ps %>% 
  select(-group) %>% 
  mutate(across(location, ~str_replace(., "\\n", " "))) %>% 
  pivot_wider(names_from = name, 
              values_from = c(mean, lower, upper)) %>% 
  mutate(across(ends_with("prob"), ~dezero(., 2)),
         across(ends_with("delta"), ~as.character(scales::comma(round(.))))) %>% 
  pivot_longer(-c(xn, location), 
               names_to = c(".value", "param"),
               names_sep = "_") %>% 
  mutate(est = str_c(mean, " [", lower, ", ", upper,"]")) %>% 
  select(-mean:-upper) %>% 
  pivot_wider(names_from = xn, 
              values_from = est)

tmp <- read_csv(file) %>% 
  filter(param %in% c("delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  pivot_wider(names_from = xn, values_from = value) %>% 
  unnest(c(masked, unmasked)) %>% 
  mutate(diff = masked - unmasked) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, recode, theta = "prob")) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location) %>% 
  select(param, 
         `Transition location` = location, 
         Unmasked = unmasked,
         Masked = masked,
         "Difference" = mean, BF)

write_csv(tmp, "tables/masking_effect_cellmeans_constrained.csv")

tmp[,-1] %>% knitr::kable(align = c("l", "r", "r", "r", "r"), 
                          caption = "Mixture model estimates for key transitions. Cell means are shown for the masked and unmasked writing task in msecs for the slowdown for long transitions and the probability of disfluent transitions. The effect for masking is shown on log scale (for transition durations) and logit scale for probability of disfluent transitions. 95% PIs in brackets.") %>%
  #  cell_spec(1, 3:4) %>% 
  kableExtra::collapse_rows(columns = 1) %>%
  kable_styling(position = "center", full_width = T) %>%
  footnote("PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.", 
           threeparttable = T) %>% 
  group_rows("Disfluencies", 1, 3) %>%
  group_rows("Probability of disfluencies", 4, 6)
